Balsa: A Compact Line Notation Based on SMILES

### 2022-07-31 manuscript.tex
Makes levels of detail consistent across intro and conclusion. Closes #44.
Uses \columnwidth for all figures. Closes #48

\begin{figure}
    \centering
    \includegraphics[width=6cm]{molecular-graph-vs-tree.pdf}
    \caption{Molecular Graph vs. Molecular Tree. The bonds in a molecular graph reference atoms indirectly through unique indexes (lower left). In contrast, a bond in a molecular tree defines its target atom inline (lower right). This increases the potential efficiency of trees over graphs for molecular representation. To simulate ring closure, a molecular tree introduces bridge nodes, as described later (solid grey circles).}
    \label{fig:molecular-graph-vs-tree}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=2.5cm]{node-order.pdf}
    \caption{Node Order. Nodes in a molecular tree assume the order in which they were added through depth-first construction. Parents precede children, and children added first preceded those added later.}
    \label{fig:node-order}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=4in]{node-and-edge-associations.pdf}
    \caption{Nodes and Edge Associations. Three node/edge associations are possible within a molecular tree. A parent atom can be connected to a child atom through a bond (left). A parent atom can also be connected to a child bridge node through a bond (center). Finally, an atom can be connected, without bonding, to a child atom through a gap (right).}
    \label{fig:nodes-and-edge-associations}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=4in]{periodic-table.pdf}
    \caption{Elements. Available chemical elements and their allowed roles. Not all IUPAC-approved elements are available. Hydrogen counts of atoms associated with elements on blue background can be computed algorithmically. Atoms associated with elements on green background can be selected.}
    \label{fig:periodic-table}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=2in]{dime.pdf}
    \caption{Delocalization-induced molecular equality (DIME). Two chemically equivalent molecular representations differ only due to delocalization.}
    \label{fig:dime}
\end{figure}

@@

This paper describes Balsa, a subset of the widely-\-used SMILES line notation. Balsa's features are detailed at four levels: constitution; delocalization; conformation; and configuration. Constitutionally, Balsa can encode any molecule conforming to the valence bond model. In the event of undesired symmetry artifacts due to delocalization, Balsa offers a mitigation based in graph theory. Conformational isomerism of alkenes is supported by partial parity bonds. The configurations of tetracoordinate, stereogenic atoms are encoded through the use of a parity enumeration and conventions around its use.

>>>>

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{molecular-graph-vs-tree.pdf}
    \caption{Molecular Graph vs. Molecular Tree. The bonds in a molecular graph reference atoms indirectly through unique indexes (lower left). In contrast, a bond in a molecular tree defines its target atom inline (lower right). This increases the potential efficiency of trees over graphs for molecular representation. To simulate ring closure, a molecular tree introduces bridge nodes, as described later (solid grey circles).}
    \label{fig:molecular-graph-vs-tree}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{node-order.pdf}
    \caption{Node Order. Nodes in a molecular tree assume the order in which they were added through depth-first construction. Parents precede children, and children added first preceded those added later.}
    \label{fig:node-order}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{node-and-edge-associations.pdf}
    \caption{Nodes and Edge Associations. Three node/edge associations are possible within a molecular tree. A parent atom can be connected to a child atom through a bond (left). A parent atom can also be connected to a child bridge node through a bond (center). Finally, an atom can be connected, without bonding, to a child atom through a gap (right).}
    \label{fig:nodes-and-edge-associations}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{periodic-table.pdf}
    \caption{Elements. Available chemical elements and their allowed roles. Not all IUPAC-approved elements are available. Hydrogen counts of atoms associated with elements on blue background can be computed algorithmically. Atoms associated with elements on green background can be selected.}
    \label{fig:periodic-table}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{dime.pdf}
    \caption{Delocalization-induced molecular equality (DIME). Two chemically equivalent molecular representations differ only due to delocalization.}
    \label{fig:dime}
\end{figure}

@@

This paper describes Balsa, a subset of the widely-\-used SMILES line notation. Balsa's features are detailed at three levels: constitution; conformation; and configuration. Constitutionally, Balsa can encode any molecule conforming to the valence bond model. In the event of undesired symmetry artifacts due to delocalization, Balsa offers a mitigation based in graph theory. Conformational isomerism of alkenes is supported by partial parity bonds. The configurations of tetracoordinate, stereogenic atoms are encoded through the use of a parity enumeration and conventions around its use.
